### 2020-05-01 Documentation/Manual.tex
Updated my affiliation and the acknowledgments (included my DOE grant)
Finally (!) updated the manual.

\newcommand{\ie}{{\it i.e.}}
\newcommand{\Ie}{{\it I.e.}}
\newcommand{\eg}{{\it e.g.}}
\newcommand{\Eg}{{\it E.g.}}
\newcommand{\cf}{{\it cf.}}
\newcommand{\etc}{{\it etc.}}
\newcommand{\eq}{Eq.}
\newcommand{\eqs}{Eqs.}
\newcommand{\Def}{Definition}
\newcommand{\fig}{Fig.}
\newcommand{\Fig}{Fig.}
\newcommand{\figs}{Figs.}
\newcommand{\Figs}{Figs.}
\newcommand{\Ref}{Ref.}
\newcommand{\Refs}{Refs.}
\newcommand{\Sec}{Sec.}
\newcommand{\Secs}{Secs.}
\newcommand{\Chapt}{Chapter}
\newcommand{\Chapts}{Chapters}
\newcommand{\Part}{Part}
\newcommand{\App}{Appendix}
\newcommand{\Apps}{Appendices}
\newcommand{\Tab}{Table}
\newcommand{\Tabs}{Tables}

@@

%\vspace*{1.5cm}
%{\large \bf !!!DRAFT IN PROGRESS!!! }
\end{center}

@@

\begin{center}
{\large Patrick Huber$^a$, Joachim Kopp$^b$, Manfred Lindner$^b$, \\ Mark Rolinec$^c$, Walter Winter$^d$}
\end{center}

@@

\begin{center}
Version from \today\ for \GLOBES\ 3.1
\end{center}

@@

{\em\small $^{\mathrm{a}}$%
       University of Wisconsin,  Physics Department,
       1150 University Av., Madison, WI 53706, USA}

{\em\small $^{\mathrm{b}}$%
       Max--Planck--Institut f\"ur Kernphysik, 
       Postfach 10~39~80, D--69029 Heidelberg, Germany}

{\em\small $^{\mathrm{c}}$%
       Technische Universit\"at M\"unchen,
       Institut f\"ur Theoretische Physik, Physik--Department,\\
       James--Franck--Strasse, D--85748 Garching, Germany}

{\em\small $^{\mathrm{d}}$%
       Universit\"at W\"urzburg, 
       Lehrstuhl f\"ur theoretische Physik II, \\
       Institut f\"ur theoretische Physik und Astrophysik, 
       Am Hubland,
       D-97074 W\"urzburg, Germany}

@@

\clearpage
\thispagestyle{empty}
\bigskip
\begin{quote}
    Copyright \copyright  2004--2010  The GLoBES Team.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.2
    or any later version published by the Free Software Foundation;
    with the invariant Sections ``Terms of usage of \GLOBES" 
    and ``Acknowledgments'', no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled "GNU
    Free Documentation License".
\end{quote}
\bigskip

@@

@@

This work and the development of \GLOBES\ have over the years been supported by
(in chronological order):
\begin{itemize}
\item Technische Universit\"at M\"unchen [All authors]
\item Max-Planck-Institut f\"ur Physik, M\"unchen [PH]
\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\
der Deutschen Forschungsgemeinschaft [All authors]
\item Studienstiftung des Deutschen Volkes [JK, WW]
\item Institute for Advanced Study, Princeton [WW]
\item W. M. Keck Foundation [WW]
\item National Science Foundation [WW]
\item University of Wisconsin, Madison [PH]
\item Max-Planck-Institut f\"ur Kernphysik, Heidelberg [JK, ML]
\item Emmy Noether-Programm der Deutschen Forschungsgemeinschaft [WW]
\item Universit\"at W\"urzburg [WW]
\end{itemize}

>>>>

\newcommand{\ie}{{\it i.e.}}
\newcommand{\Ie}{{\it I.e.}}
\newcommand{\eg}{{\it e.g.}}
\newcommand{\Eg}{{\it E.g.}}
\newcommand{\cf}{{\it cf.}}
\newcommand{\etc}{{\it etc.}}
\newcommand{\eq}{Eq.}
\newcommand{\eqs}{Eqs.}
\newcommand{\Def}{Definition}
\newcommand{\fig}{Fig.}
\newcommand{\Fig}{Fig.}
\newcommand{\figs}{Figs.}
\newcommand{\Figs}{Figs.}
\renewcommand{\Ref}{Ref.}
\newcommand{\Refs}{Refs.}
\newcommand{\Sec}{Sec.}
\newcommand{\Secs}{Secs.}
\newcommand{\Chapt}{Chapter}
\newcommand{\Chapts}{Chapters}
\newcommand{\Part}{Part}
\newcommand{\App}{Appendix}
\newcommand{\Apps}{Appendices}
\newcommand{\Tab}{Table}
\newcommand{\Tabs}{Tables}

@@

\vspace*{1.5cm}
{\large \bf !!!DRAFT IN PROGRESS!!! }
\end{center}

@@

\begin{center}
{\large Patrick Huber$^a$, Joachim Kopp$^b$, Manfred Lindner$^c$, \\ Walter Winter$^d$}
\end{center}

@@

\begin{center}
Version from \today\ for \GLOBES\ 3.2.18
\end{center}

@@

{\em\small $^{\mathrm{a}}$%
       Virginia Tech, Center for Neutrino Physics, Blacksburg, VA 24061, USA}

{\em\small $^{\mathrm{b}}$%
  Theoretical Physics Department,
  CERN and Johannes Gutenberg University Mainz (Germany),
  1211 Geneve 23, Switzerland
}

{\em\small $^{\mathrm{c}}$%
       Max--Planck--Institut f\"ur Kernphysik, 
       Postfach 10~39~80, D--69029 Heidelberg, Germany}

{\em\small $^{\mathrm{d}}$%
Deutsches Elektronen-Synchrotron (DESY)
Platanenallee 6
15738 Zeuthen, Germany}

@@

\clearpage
\thispagestyle{empty}
\bigskip
\begin{quote}
    Copyright \copyright  2004--2020  The GLoBES Team.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.2
    or any later version published by the Free Software Foundation;
    with the invariant Sections ``Terms of usage of \GLOBES" 
    and ``Acknowledgments'', no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled "GNU
    Free Documentation License".
\end{quote}
\bigskip

@@

\section*{Version 3.2.18}
\index{norm}{Version 3.2.18}

\begin{itemize}

  \item Systematics correlated across experiments/detector; see
    \Sec~\ref{sec:newsys}. This feature has been extensively tested
    and used, amongst others, for the DUNE CDR~\cite{Acciarri:2015uup}.
\item Fitting of actual data; see \Sec~\ref{sec:fitting}.
\end{itemize}

@@

This work and the development of \GLOBES\ have over the years been supported by
(in chronological order):
\begin{itemize}
\item Technische Universit\"at M\"unchen [All authors]
\item Max-Planck-Institut f\"ur Physik, M\"unchen [PH]
\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\
der Deutschen Forschungsgemeinschaft [All authors]
\item Studienstiftung des Deutschen Volkes [JK, WW]
\item Institute for Advanced Study, Princeton [WW]
\item W. M. Keck Foundation [WW]
\item National Science Foundation [WW]
\item University of Wisconsin, Madison [PH]
\item Max-Planck-Institut f\"ur Kernphysik, Heidelberg [JK, ML]
\item Emmy Noether-Programm der Deutschen Forschungsgemeinschaft [WW]
\item Universit\"at W\"urzburg [WW]
\item US Department of Energy under award number \protect{DE-SC0003915} [PH]
\end{itemize}

### 2020-05-01 Documentation/part1.tex
Finally (!) updated the manual.

>>>>

A related function which allows to name an oscillation engine without invoking it
\begin{function}
  \GLBNS{DefineOscEngine}
{\tt int glbDefineOscillationEngine(int n\_parameters,
                 glb\_probability\_matrix\_function prob\_func, \\
                 glb\_set\_oscillation\_parameters\_function set\_params\_func, \\
                 glb\_get\_oscillation\_parameters\_function get\_params\_func, \\
                 const char[] name,\\
                 void* user\_data)} defines a probability
engine for the simulation of non-standard physics with {\tt n\_parameters} oscillation parameters ({\tt n\_parameters} $\ge6$). The three functions {\tt prob\_func}, {\tt set\_params\_func}, and {\tt get\_params\_func} 
will be used with the function types defined above. In order to circumvent global variables, an arbitrary pointer {\tt user\_data} can be defined, that will be passed to the probability engine in each function call. {\tt name} can be use to refer to the so defined oscillation engine within {\sf AEDL} using the {\tt \$oscillation\_engine} directive.
\end{function}
\index{aedl}{oscillation engine@{\tt "\$oscillation\_engine}}

### 2020-05-01 Documentation/part2.tex
Further functions added to manual
Finally (!) updated the manual.

We now give a short introduction to the syntax of \AEDL .
 The first eight characters have to be {\tt \%!GLoBES}
in order to avoid parsing megabytes of chunk
 and producing thousands of error messages.  In addition, the minimum
\GLOBES\ version that the \AEDL\ file is supposed to run with has to be
defined by a {\tt \$version} statement, such as
{\tt \$version="3.0.0"}.
\index{aedl}{version@{\tt \$version}}
%
Comments can be used in the same way as in C:
\begin{quote}
{\tt /* This starts a comment\\
 and here the comment ends */ \\
// Another comment
}
\end{quote}
There are pre-defined variables which all start with {\tt \$}. Their range
is also checked. For example,  {\tt 
\$bins} can be only between $0$ and $500$.\footnote{The upper limit is 
only there for safety reasons, the memory is allocated dynamically.} If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.  For example, we have scalar variables
\begin{quote}
{\tt
\$bins = 10\\
\$baseline = 1200.0
}
\end{quote}
and simple lists
\begin{quote}
{\tt
\$densitytab=\{1.0,2.2343,3.3432\} 
}
\end{quote}
%
Since there are often groups of data which we want to refer to later,
environments can be used. This is illustrated 
with the channel definition part:
\begin{quote}
{\tt channel(\#ch1)<\\
\tb  $\ldots$\\
>
}
\end{quote}
The first part is the type of environment, which is {\tt channel} here. 
There are the following types of environments in \AEDL :
\begin{quote}
{\tt nuflux\\
cross\\
channel\\
energy\\
rule
}
\end{quote}
Besides the environment type, there is a user-defined name 
beginning with {\tt \#}
in the above example: {\tt \#ch1}. It can be used later to refer to the 
channel defined in {\tt <$\ldots$>}. Those names are so-called 
``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be referred to after their definition.
However, similar to C, one can give a declaration without definition before:
\begin{quote}
{\tt    channel(\#ch2)<>}
\end{quote}
Now one can refer to the name {\tt \#ch2}, while the actual channel definition comes later. The internal representation of this automatic
variable is a number, which obtains its value from a counter for each type of environment. For example, for {\tt channel} the counter is {\tt numofchannels}. The counter keeps track of how many different names 
there are for one type of environment, which means that it counts the number of channels, rules, energy resolution functions, \etc . Thus, the automatic
variables are numbered in the order of their definition, and the number
can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). In order to facilitate the the mapping from names in \AEDL\ to indices
in C there are two functions \GLB{NameToValue} and \GLB{ValueToName} which
make this transition (see \Sec~\ref{sec:aedl_names}, 
page~\pageref{sec:aedl_names}).

@@

\index{aedl}{NEXT@{\tt \#NEXT\#}}
If you want to have several experiments in one file, separate the different
 experiments by 
\begin{quote}
{\tt    \#NEXT\#}
\end{quote}
This command resets the counters for channels, rules, fluxes, cross section 
and energy resolution environments. All variables have their scope limited 
by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}.  This allows 
a consistent treatment of various experiments in one file.

@@

>>>>

We now give a short introduction to the syntax of \AEDL .
 The first eight characters have to be {\tt \%!GLoBES}
in order to avoid parsing megabytes of chunk
 and producing thousands of error messages.  The minimum
\GLOBES\ version that the \AEDL\ file is supposed to run with has to be
defined by a {\tt \$version} statement, such as
{\tt \$version="3.0.0"}.
\index{aedl}{version@{\tt \$version}} 
Furthermore, references linked to the \AEDL\ file can be defined with {\tt \$citation="..."},
which can later be retrieved with {\tt glbGetCitationForExperiment}, see page~\pageref{glbGetCitationForExperiment}.\footnote{Available in \GLOBES\ 3.1.7 and higher.}
\index{aedl}{citation@{\tt \$citation}} \label{citation}
%
Comments can be used in the same way as in C:
\begin{quote}
{\tt /* This starts a comment\\
 and here the comment ends */ \\
// Another comment
}
\end{quote}
There are pre-defined variables which all start with {\tt \$}. Their range
is also checked. For example,  {\tt 
\$bins} can be only between $0$ and $500$.\footnote{The upper limit is 
only there for safety reasons, the memory is allocated dynamically.} If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.  For example, we have scalar variables
\begin{quote}
{\tt
\$bins = 10\\
\$baseline = 1200.0
}
\end{quote}
and simple lists
\begin{quote}
{\tt
\$densitytab=\{1.0,2.2343,3.3432\} 
}
\end{quote}
%
Since there are often groups of data which we want to refer to later,
environments can be used. This is illustrated 
with the channel definition part:
\begin{quote}
{\tt channel(\#ch1)<\\
\tb  $\ldots$\\
>
}
\end{quote}
The first part is the type of environment, which is {\tt channel} here. 
There are the following types of environments in \AEDL :
\begin{quote}
{\tt nuflux\\
cross\\
channel\\
energy\\
rule
}
\end{quote}
Besides the environment type, there is a user-defined name 
beginning with {\tt \#}
in the above example: {\tt \#ch1}. It can be used later to refer to the 
channel defined in {\tt <$\ldots$>}. Those names are so-called 
``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be referred to after their definition.
However, similar to C, one can give a declaration without definition before:
\begin{quote}
{\tt    channel(\#ch2)<>}
\end{quote}
Now one can refer to the name {\tt \#ch2}, while the actual channel definition comes later. The internal representation of this automatic
variable is a number, which obtains its value from a counter for each type of environment. For example, for {\tt channel} the counter is {\tt numofchannels}. The counter keeps track of how many different names 
there are for one type of environment, which means that it counts the number of channels, rules, energy resolution functions, \etc . Thus, the automatic
variables are numbered in the order of their definition, and the number
can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). In order to facilitate the the mapping from names in \AEDL\ to indices
in C there are two functions \GLB{NameToValue} and \GLB{ValueToName} which
make this transition (see \Sec~\ref{sec:aedl_names}, 
page~\pageref{sec:aedl_names}).

@@

\index{aedl}{NEXT@{\tt \#NEXT\#}}
If you want to have several experiments in one file, separate the different
 experiments by 
\begin{quote}
{\tt    \#NEXT\#}
\end{quote}
This command resets the counters for channels, rules, fluxes, cross
section and energy resolution environments. All, except named nuisance
parameters (see. Sec.~\ref{sec:newsys}), variables have their scope limited by either {\tt
  \%!GLoBES, \#NEXT\#} or {\tt EOF}.  This allows a consistent
treatment of various experiments in one file.

\index{aedl}{DETECTOR@{\tt \#DETECTOR\#}}
If you want to have several detectors in the same beam, separate the different
 detectors by 
\begin{quote}
{\tt    \#DETECTOR\#}
\end{quote}
This command resets the counters and names \emph{only} for rules. All
other variables have global scope.  This feature is useful to define
correlated systematics for near and far detector type setups as
described in Sec.~\ref{sec:newsys}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Systematics correlated accross experiments and rules*}
\label{sec:newsys}

\index{norm}{Systematics!correlated|(} 
\index{aedl}{sys@{\tt sys}|(} In many cases the simple rule-based
style of systematics is sufficient, but there are cases, notably DUNE
and almost all modern reactor experiments, where a more fine-grained
systematics implementation is necessary.  The way to accomplish is to introduce named nuisance parameters:
\begin{quote}
  {\tt sys(\#norm)<\\
    \tb @systype=0\\
\tb  @error = 0.02\\
>}
\end{quote}

\index{aedl}{sys@{\tt sys}!error@{\tt "@error}}

\index{aedl}{sys@{\tt sys}!systype@{\tt "@systype}}

{\tt @systype} defaults to $0$, a common multiplicative nuisance
parameter.  {\tt @systype} can also be set to $1$ where this becomes a
energy calibration error as desribed in Sec.~\ref{sec:energy}.

In this case we introduced a nuisance parameter {\tt \#norm} with an
error of $0.02$ and a central value of $1$. These named nuisance
parameters now can be used in the definition of a rule:

\begin{quote}
  {\tt rule(\#NU\_E\_Disappearance)<\\
\tb    @signal     = 0.4@\#nu\_e\_disappearance\_CC\_U235 \\
\tb    @sys\_on\_multiex\_errors\_sig =\{\#norm,\#flux235\} \\
\tb    @background = 0.1@\#nu\_e\_bg\\
\tb     @sys\_on\_multiex\_errors\_bg  = {\#bgnorm}\\
\tb        @sys\_on\_function  = "chiMultiExp"\\
\tb        @sys\_off\_function = "chiNoSysSpectrum"\\
>}
\end{quote}

The key here is the line {\tt @sys\_on\_multiex\_errors\_sig
  =\{\#norm,\#flux235\} }, where the tag {\tt
  @sys\_on\_multiex\_errors\_sig } is invoked to signal the use of
this feature and each {\tt \{ ...\}} contains all named nuisance
parameters applicable to the corresponding part of {\tt @signal}
definition; the same syntax applies to {\tt
  @sys\_on\_multiex\_errors\_bg } for the {\tt @background} part of
the rule.  {\tt @sys\_on\_function = "chiMultiExp"} ensures that the
appropriate $\chi^2$-function is used to handle named nuisance
parameters.%

\index{aedl}{rule@{\tt rule}!sys on multiex errors  sig@{\tt "@sys\_on\_multiex\_errors\_sig}}

\index{aedl}{rule@{\tt rule}!sys on multiex errors bg@{\tt "@sys\_on\_multiex\_errors\_bg}}

Now this very same technique can be used in different rules of the
same experiment but also in rules defined in \emph{different}
experiments. All instance of the named parameter, in this example {\tt
  \#norm}, will be fully correlated. In the context of
near/far-detector experiments also the {\tt \#DETECTOR\#} directive is
interesting since it allows to reuse flux, channel etc. definitions
seamlessly. It is important that all named parameters have been
defined using the {\tt sys<>} environment before they are used. It is
good practive to put all {\tt sys<>} definitions in one file and to
include this file into all subsequent {\sf AEDL} files.

Sometimes, systematics are energy dependent and this can be implemented within the same framework as
\begin{quote}
  {\tt sys(\#norm)<\\
    \tb @energy\_list =\{ 0.002,0.005,0.006,0.009\}\\
    \tb  @error\_list = \{0.02,0.01,0.005,0\}\\
>}
\end{quote}
\index{aedl}{sys@{\tt sys}!error list@{\tt "@error\_list}}
\index{aedl}{sys@{\tt sys}!energy list@{\tt "@energy\_list}}

This effectively introduces 1 nuisance parameter in the fit which
varies the rate by 0.02 at 0.002\,GeV and by 0.01 at 0.005\,GeV
asf. fully correlated among all energies. For bin center energies
between the support points in {\tt @energy\_list} linear interpolation is
used. Both lists need to have the same number of elements.

\index{aedl}{sys@{\tt sys}|)}
\index{norm}{Systematics!correlated|)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fitting actual experimental data*}
\label{sec:fitting}

\index{norm}{Fitting of data|(}
The framework presented so far is general enough to achieve an
accurate decsription of real experiments and thus, the user may want
to fit actual data with {\sf GLoBES}. This can be conveniently achieved by using

\begin{quote}
  {\tt rule(\#foo)<\\
    \tb all the usual definitions of a rule\\
    \\
    \tb @data\_flag=1\\
    \tb @data=\{1.23,2.34,3.45,...,42\}\\
>}
\end{quote}
\index{aedl}{rule@{\tt rule}!data flag@{\tt "@data\_flag}}
\index{aedl}{rule@{\tt rule}!data@{\tt "@data}}

where {\tt @data} is a list of the actual data which needs to have
{\tt \$bins} elements and the {\tt @data\_flag} can be use to turn
fitting of this data on (1) or off (0).

In some cases a non-standard oscillation probability calculaton may be
required, to include for instance the size of the neutrino production
region. This requires to use the mechanism to define an oscillation
engine and to name it, this is done with {\tt DefineOscEngine} and to
then call the corresponding name from withing {\sf AEDL}, which is
accomplished with the {\tt \$oscillation\_engine} directive. Note,
that the scope of {\tt \$oscillation\_engine} is not limited by {\tt
  \#DETECTOR\#}.

\GLBNS{DefineOscEngine}
\index{aedl}{oscillation engine@{\tt "\$oscillation\_engine}}

\index{norm}{Fitting of data|)}
